### 2015-04-27 Paper_v1.tex
Updated table

The objective function in Equation 1 is minimized using appropriate optimization technique to estimate the model parameters. These problems are generally multi-modal i.e. contain multiple local optima traditional. Hence we cannot use local optimization techniques cannot be used to solve these problems. In many of these problems derivative information is absent or the objective function may be discontinuous. This hinders the use of deterministic optimization techniques that require the problem to be convex and other derivative/gradient based optimization methods. Most of these methods are computationally expensive and for optimization of large-scale ODE systems, this becomes prohibitively expensive. Heuristic optimization techniques like genetic algorithm and differential evolution offer an attractive option in finding globally optimum solution or close to global optimum solution. However most of the population-based heuristics take a large number of function evaluations and have computationally complex operations, which reduces their effectiveness in finding feasible solutions with limited computational resources and time.

This study tries to overcome the drawbacks mentioned above and presents a novel heuristics based search strategy/algorithm for parameter estimation in large-scale ODE models of biochemical networks with many unknown parameters.

@@

\subsection *{Modeling methodologies} 
 Shuler and co-workers constructed large-scale dynamic metabolic models of E.coli that were based ordinary differential equations (ODEs). Various aspects of E.colis cellular processes like RNA synthesis, chromosome synthesis, cellular energy expenditure etc. were modeled using coupled non-linear ordinary differential equations. Post genomic era saw the rise of flux balance analysis (FBA), a static, constraint based modeling approach for microbial metabolism. FBA assumes a pseudo steady-state and reduces genome-scale kinetic models to a set of static mass balances. These mass balances are represented using an underdetermined set of linear algebraic equations. The advantage is such a system of equations can be solved efficiently for very large systems.

@@

\section*{Results}\vspace{-12pt}

@@

Villaverde and co-workers recently published a set of benchmark problems to evaluate parameter estimation methods [REFHERE]. From a computational cost perspective problems they categorized the problems as most expensive, intermediate and least expensive. We evaluated the performance of our algorithms on a problem from each category. Table zzz shows that

\begin{table}[h!]
\centering
\begin{tabular}{|c c c c|} 
 \hline
 ModelID & B1 & B4&\\ [0.5ex] 
 \hline
 Upper bound & ${5.pnom}$ & ${5.pnom}$ & \\  [0.5ex]
 Lower bound & ${0.2.pnom}$ & ${0.2.pnom}$ &\\ [0.5ex]
 CPU time & $38.308$ ${hours}$ & $6.2$ ${minutes}$& \\ [0.5ex]
 Function Evaluations & $4000$ & $4000$&\\ [0.5ex]
 Initial Objective Value & $1.0589.10 ^{10}$ & $1.4275.10 ^{7}$ &\\ [0.5ex] 
 Final Objective Value & $4.96.10^{5}$ & $38.9375$&\\ [0.5ex]
 Nominal Objective Value & $1.0986·10^{6}$ & $39.0676$&\\ [1ex]
 \hline
\end{tabular}
\caption{Table to test captions and labels}
\label{table:1}
\end{table}

>>>>

The objective function from equation ---  is minimized using appropriate optimization technique to estimate the model parameters. Due to the complex dynamics of biological systems these problems are multi-modal i.e. most of them have multiple local minima or maxima. The non-linearity of the problem coupled with multi-modality is a substantial challenge and generally renders the local optimization techniques incapable of obtaining an optimal solution.   
\textit {In many of these problems derivative information is absent or the objective function may be discontinuous. This hinders the use of deterministic optimization techniques that require the problem to be convex and other derivative/gradient based optimization methods. Most of these methods are computationally expensive and for optimization of large-scale systems, this becomes prohibitively expensive.  Heuristic approaches have been promising in this regard, however most of the population-based heuristics take a large number of function evaluations and have computationally complex operations, which reduces their effectiveness in finding feasible solutions with limited computational resources and time. Hybrid methods that couple heuristics with derivative based methods}

The problem get further compounded when we have parameter vectors of very high dimensions. In this study we developed a novel meta heuristic approach that combines multi-swarm particle swarm optimization with dynamically dimensioned search (DDS)to obtain optimal or near optimal solutions to high-dimensional, non-linear biological models within a relatively few function evaluations.

A bit of history on DDS
Why coagulation model? feed back loops, parameter identifiability, tight regulation and also high dimensional

@@

\subsection *{Modeling methodologies Literature} 
 Shuler and co-workers constructed large-scale dynamic metabolic models of E.coli that were based ordinary differential equations (ODEs). Various aspects of E.colis cellular processes like RNA synthesis, chromosome synthesis, cellular energy expenditure etc. were modeled using coupled non-linear ordinary differential equations. Post genomic era saw the rise of flux balance analysis (FBA), a static, constraint based modeling approach for microbial metabolism. FBA assumes a pseudo steady-state and reduces genome-scale kinetic models to a set of static mass balances. These mass balances are represented using an underdetermined set of linear algebraic equations. The advantage is such a system of equations can be solved efficiently for very large systems.

@@

\section*{Results}
\subsection{Coagulation model formulation}

@@

Villaverde and co-workers recently published a set of benchmark problems to evaluate parameter estimation methods [REFHERE]. From a computational cost perspective problems they categorized the problems as most expensive, intermediate and least expensive. We evaluated the performance of our algorithms on a problem from the most expensive and least expensive categories. Table 1 shows that

\begin{table}[h!]
\centering
\begin{tabular}{|c c c c c c c|} 
 \hline
 ModelID & Coagulation & B1 & B4 & Ackley300D & Rastrigin300D&\\ [0.5ex] 
 \hline
 Upper bound&$1000.pnom$& ${5.pnom}$ & ${5.pnom}$ &${30}$ &$5.12$& \\  [0.5ex]
 Lower bound&$0.001.pnom$& ${0.2.pnom}$ & ${0.2.pnom}$&${-15}$&$-5.12$&\\ [0.5ex]
 CPU time &$5.6$ ${hours}$ &$38.308$ ${hours}$ & $6.2$ ${minutes}$& ${2.8}$${seconds}$&${2.6}$${seconds}$& \\ [0.5ex]
 Function Evaluations &$4000$&$4000$&$4000$&$4000$&$4000$&\\ [0.5ex]
 Initial Objective Value & $1.0589.10 ^{10}$ & $1.4275.10 ^{7}$ &$1.8536.10 ^{7}$&$21.12$&$99.985$&\\ [0.5ex] 
 Final Objective Value & $2.9223.10^{3}$ & $4.96.10^{5}$ & $38.9375$&$0$&$0$&\\ [0.5ex]
 Nominal Objective Value & $4.7785·10^{6}$& $1.0986·10^{6}$ & $39.0676$&$0$&$0$&\\ [1ex]
 \hline
\end{tabular}
\caption{Table with optimization settings and results for coagulation and different benchmarks using DDSMLSPSO.}
\label{table:1}
\end{table}
